% Appendix A

\chapter{Practical Guidance} % Main appendix title

\label{AppendixA} % For referencing this appendix elsewhere, use \ref{AppendixA}

\section{Navigating The Ethereum Ecosystem}
The Ethereum ecosystem is relatively busy place and can be disorientating for newcomers. This section serves as a short glossary to make this process a little easier.\\
\\
\keyword{geth}\\
Geth is a full Ethereum client written in the go language. It implements the Web3 Javascript API over JSON RPC.\\
 \\
\keyword{web3.js}\\
A javascript library that implements needed to connect to a local Ethereum client.\\
\\
\keyword{TurboEthereum}\\
The name for the software distribution implemented in C++. It?s major components are:\\
\\
\begin{itemize}
\item AlethOne. A 'Mainline' Ethereum desktop miner
\item AlethZero. A 'Power-User' Ethereum client.
\item ++eth. The 'mainline' CLI Ethereum client. 
\item Mix. An IDE (integrated Development Environment) for building dapps. 
\item TurboEthereum (C++ Ethereum Client)
\end{itemize}
\keyword{Webthree-Umbrella}\\
This is essentially just the name of the umbrella repository for TurboEthereum.\\
\\
\keyword{Pyeth}\\
Pyeth is a Python implementation of an Ethereum client. There are a number of clients in different languages.\\
\\
\keyword{Solidity, Serpent \& Mutan}\\
Languages for smart contract development. New developers are advised to stick to Solidity.\\
\\
\keyword{Mist}\\
Mist is a user friendly browser for Ethereum dapps. The Mist browser will serve as a wallet and interface for dapps. It is scheduled to be launched at the Metropolis release stage.\\
\\
\keyword{Casper}\\
Casper is the proof-of-stake consensus protocol planned to be launched with the Serenity Ethereum release. It is so called because it is an adaption of the GHOST (Greedy Heaviest Observed Subtree First) protocol.\\

\section{Creating a Private Test Network}
A local private network is an Ethereum network with one peer or miner which is your host machine. To set up a local instance you will need to stop your client from attaching to existing networks and 'listening for peers'. To make development a bit easier it is suggested that a custom genesis block is used with a decreased difficulty and some pre-seeded accounts. Ade Duke has written a good blog post on 'How To Create A Private Ethereum Chain'\cite{AdeDuke}, more practical information can be found there.\\





